---
title: "Data Management - Part B"
author: "GROUP_12"
output:
  word_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---
## Set Up Environment

```{r setup,attr.source='.numberLines'}
#set up the environment
library("dplyr")
library("tidyr")
library("tidyverse")
knitr::opts_chunk$set(echo = TRUE)
```

## First Step: Get the Directory of Each .xlsx

```{r,attr.source='.numberLines'}
# get the directory of each .xlsx of each country
dic<-list.files("group_assignment",recursive = TRUE)

# delete the unwanted .xlsx in directory
dic1<-dic[-1]

# prepare an empty list to save the directory later
dic2 <- list()

# set count to 0 for running the for loop to select the number of the list later
count<-0

# combine the root directory and .xlsx directory into the list
for (f in 1:length(dic1)){
  # every time one new directory is dene, count+1 to move to next space of list
  count<-count+1
  # select the space of list and put the combined string in it
  dic2[count]<-list(paste0("group_assignment/",dic1[f]))
}
```

## Second Step: Import, Clean, and Combine Each .xlsx

```{r message = FALSE,attr.source='.numberLines'}
# create an empty data frame for combining all clean .xlsx in for loop later
df <- data.frame()

# set count2 to 0 for running the for loop to read the .xlsx in order later
count2<-0

# use the for loop to deal with .xlsx one by one
for (f in 1:length(dic2)) {
  # at the begining, increase one in count2 to specify which number of .xlsx we're reading in the list dic2
  count2<-count2+1
  # use read_xlsx function to read the .xlsx document, and also use paste function to transfer the list into string.
  # we skip firt row in each .xlsx to avoid loading the url data, which is useless in this part
  data <- readxl::read_xlsx(paste(dic2[count2]),skip=1)
  # save the country data into a new variable before we clean the redundant data
  country <- data$...3[3]
  # same purpose as country to flow
  flow <- data$...3[1]
  # clean the redundant data in the file
  data <- data[-c(1:3,5,(nrow(data)-2):(nrow(data))),-2]
  # save the first row of data into col.name.list to replcae the columns name later
  col.name.list<-data[1,]
  # replace the wrong columns name with first row of data, which is saved in the col.name.list
  colnames(data)<-col.name.list
  # remove the first row of data after changing it to columns name
  data <- data[-1,]
  # transform the wide file into long file
  data_gather<-gather(data,"product","value",-"Product")%>%
  # correct the name of year column
      rename(year=Product)
  # add a new column to put the country inside the table
  data_gather$country<-country
  # add a new column to put the flow inside the table
  data_gather$flow<-flow
  # put the clean and structured table into the data frame
  df <- rbind(df,data_gather)
}
```

## Third Step: Arrange and Finalise Data

```{r,attr.source='.numberLines'}
# rearrange columns
data_final <- df[,c(4,1,5,2,3)]

# change .. to NA
data_final$value[which(data_final$value=="..")]<-NA

# arrange the data as requirement
data_final <- data_final%>%
  arrange(country,flow,year)

# show the final data
print(data_final)
```

We keep the NA without omitting or replacing with 0 because we assume that removing NA results in removing some data, which is not a good thing for managing data, and that NA is different from 0. However, NA means there is no record for that product for that country in that year. Hence, when we count the rows of records, we will count without NA.

```{r,attr.source='.numberLines'}
# For the total number of records on the dataset without na
row_number_without_na <- nrow(data_final%>%na.omit())
print(row_number_without_na)

# For the total number of records for each product across countries across years without na
count_product_without_na <- data_final%>%
  na.omit()%>%
  group_by(product,country,year)%>%
  summarise(n = n())%>%
  arrange(desc(n))
print(count_product_without_na)
```
